### 2023-12-12 main.tex
ss
ll

where $\mathcal{P}_{x_1}(1)$ is the probability that a single random walk started at $x_1$ is absorbed {\color{blue}at site $0$}.\todo{site 0?} Then, using \eqref{exactCorrelations}, we have

@@

\todo{here Y instead of $\tau$? [DONE]}

@@

Therefore, we compute the third cumulant, i.e. the three-point connected correlation as 
\begin{align}
	&\mathbb{E}\left[\left(\mathbbm{1}_{\{Y_{x_{1}}=a_{1}\}}-\langle\rho_{a_{1}}^{x_{1}}\rangle \right)\left(\mathbbm{1}_{\{Y_{x_{2}}=a_{2}\}}-\langle\rho_{a_{2}}^{x_{2}}\rangle \right)\left(\mathbbm{1}_{\{Y_{x_{3}}=a_{3}\}}-\langle\rho_{a_{3}}^{x_{3}}\rangle \right)\right]\nonumber\\
	=&-\frac{2x_{1}(L+1-2x_{2})(L+1-x_{3})}{(L+1)^{3}(L-1)L}(\alpha_{a_{1}}-\beta_{a_{1}})(\alpha_{a_{2}}-\beta_{a_{2}})(\alpha_{a_{3}}-\beta_{a_{3}})\dt
\end{align}
\todo{should this all be $Y$ in the identities?[DONE]}
\begin{remark} The first and second cumulants, computed in \eqref{one-pts-corr} and \eqref{two-pts-corr} respectively, match with those found in Section 4.3 of \cite{vanicat2017exact}.
	\end{remark}

@@

	\todo{added ref... What about Section 4.4.5?[PREVIOUS SECTION 4.4.5 NOW MOVED IN THE APPENDIX A]}

@@

\paragraph{Proof of formula \eqref{ResulsBasis}}  
We consider the vector with elements given by the operators of the MPA $(X_{1},\ldots,X_{N})$ and we act on it with the similarity transformation $S=S_{2}S_{1}$ obtaining new operators $(\Xt_{1},\ldots,\Xt_{N})$ that will satisfy simpler commutation relations. 
% For the sake of notation we denote by
% \begin{equation}
%     \begin{pmatrix}
% 		X_{a}\\ 
% 		X_{N}
% 	\end{pmatrix}=\begin{pmatrix}
% 	    X_{1}\\
%      \vdots\\
%      X_{N}
% 	\end{pmatrix}\co
% \end{equation}
We define the transformed matrix product operators
\begin{equation}\label{Xtildes2b}
	\begin{pmatrix}
		\Xt_{1}\\ 
		\vdots\\
		\Xt_{N-1}\\
		\Xt_{N}
	\end{pmatrix} =S\begin{pmatrix}
		X_{1}\\
		\vdots 
		\\
		X_{N-1}
		\\X_{N}
	\end{pmatrix}=\begin{pmatrix} 
		X_{1}-\beta_{1}(X_{1}+\ldots+X_{N})\\
		\vdots \\
		X_{N-1}-\beta_{N-1}(X_{1}+\ldots+X_{N})\\
				X_{1}+\ldots +X_{N}\\
	\end{pmatrix}\dt
\end{equation}
We can also reverse the transformation by $S^{-1}$ and get: 
\begin{equation}\label{Xes}
 X_{a}=	\Xt_{a}+\beta_{a}\Xt_{N}\,,\qquad
	X_{N} = \beta_N\Xt_{N}-(\Xt_{1}+\ldots+\Xt_{N-1})
\end{equation}
where  $ a\in \{1,\ldots,N-1\}$.

Summing over $b\in \{1,\ldots,N\}$ in commutation relations \eqref{bulk} we have
\begin{equation} 
	\left[X_{a},\Xt_{N}\right]=(\alpha_{a}-\beta_{a})\Xt_{N}
\end{equation}
with $\widetilde{X}_{N}= {X}_{1}+\ldots+ {X}_{N}$. 
Therefore, using \eqref{Xes}
we obtain the commutation relations for $\Xt_{a}$ 
\begin{equation}\label{commutationsBula}
	\left[\Xt_{a},\Xt_{N}\right]=(\alpha_{a}-\beta_{a})\Xt_{N}\co
\end{equation}
for $a=1,\ldots,N-1$,
cf.~\eqref{commutationsBula2}.
Moreover using \eqref{Xtildes2b} and \eqref{Xes}, the action of $\Xt_{a}$ on the boundary vectors are given by 
\begin{equation}\label{commLEFT}
	 \langle \langle W|  \left((\alpha_{a}-\beta_{a})\Xt_{N}-\Xt_{a}\right)=(\alpha_{a}-\beta_{a}) \langle \langle W|  
	\co
\end{equation}
\begin{equation}\label{commRIGHT}
	\Xt_{a}  |V\rangle\rangle = (\alpha_{a}-\beta_{a}) |V\rangle\rangle 
	\co
\end{equation} 
for $a=1,\ldots,N-1$,
cf.~\eqref{commLEFT2}.
Using the transformed operators $(\widetilde{X}_{a})_{a\in\{1,\ldots,N\}}$, the vector $|\Psi^{''}\rangle$ of the matrix product ansatz is written as
\begin{equation}\label{toDetermineTilde}
	|\Psi^{''}\rangle = \frac{1}{Z_{L}}\sum_{\bm{\tau}\in \Omega^{'}} \langle \langle W| \widetilde{X}_{\tau_{1}}\cdots \widetilde{X}_{\tau_{L}}
	|V\rangle\rangle \,|\bm{\tau}\rangle
\end{equation} 
where
the normalization is given by 
\begin{align}
		Z_{L}&= \langle \langle W|  (X_{1}+\ldots+X_{N})^{L} |V\rangle\rangle = \langle \langle W|  \widetilde{X}_{N}^{L} |V\rangle\rangle \dt
\end{align}

@@

\todo{ok? adopt notation?[DONE]}

@@

\todo{double check [DONE]}

@@

\todo{is there $c_j$ and $c_k$ missing?[DONE]}

@@

		\xi_{A}^{x}=\sum_{y=1}^{L}\sum_{b=1}^{N-1}\mathbbm{1}_{\{y=x_{k}\}}\mathbbm{1}_{\{b=a_{k}\}}\qquad &\text{if}\quad x=x_{k}\;\text{and} \; A=a_{k}\\
		\xi_{A}^{x}=0\qquad &\text{otherwise}

@@

@@

		\mathcal{L}_{\text{left}}^{TH}f(\bm{n})=&\sum_{r_{1}=0}^{\nu}\ldots\sum_{r_{N}=0}^{\nu}\mathbbm{1}_{\{r_{1}+\ldots+r_{N}=\nu\}}\left\{f(r,n^{2},\ldots,n^{L})-f(\bm{n})\right\}\\&

@@

@@

@@

\subsubsection{Duality for the reaction-diffusion process}
In this section we formulate duality for the multispecies reaction diffusion process. This answer to the second question addressed at the beginning of Section~\ref{Subsection-RD}.  The dual process is defined on the enlarged state space $\widetilde{\Omega}_{L}$.  
\begin{proposition}[Duality for the reaction-diffusion process]\label{propositin-duality-RD}
The reaction-diffusion multispecies stirring process $(\bm{n}(t))_{t\geq 0}$, on the state space $\Omega_{L}$, with generator $\mathcal{L}^{rd}$ defined in \eqref{RDGenerator} is dual to the process $(\bm{\xi}(t))_{t\geq 0}$ on the state space $\widetilde{\Omega}_{L}$ with dual generator
\begin{equation}\label{DualGeneratorRD}
	\widetilde{\mathcal{L}}_{\text{left}}^{rd}=\widetilde{\mathcal{L}}^{rd}+\sum_{x=1}^{L-1}\mathcal{L}_{x,x+1}^{rd}+\widetilde{\mathcal{L}}_{\text{right}}^{rd}
\end{equation}
where 
$\mathcal{L}_{x,x+1}^{rd}$ is defined in \eqref{edgeGeneratorRD} and, $\widetilde{\mathcal{L}}_{\text{left}}^{rd}$ and $\widetilde{\mathcal{L}}_{\text{right}}^{rd}$ are the absorbing dual generators $\widetilde{\mathcal{L}}_{x}$ defined in \eqref{siteDualGenerator} and acting at sites $1$ and $L$, with the specific choice of the reservoirs parameters \eqref{boundaryParamRD}. The duality function
is the same of the stirring process, i.e.
\begin{equation}
	D(\bm{n},\bm{\xi})=\prod_{a=1}^{N-1}\left(\rho_{a}^{\text{left}}\right)^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\frac{(\nu -\sum_{a=1}^{N-1}\xi_{a}^{x})!}{\nu!}\prod_{a=1}^{N-1}\frac{n_{a}^{x}!}{(n_{a}^{x}-\xi_{a}^{x})!}\right)\prod_{a=1}^{N-1}\left(\rho_{a}^{\text{right}}\right)^{\xi_{a}^{L+1}}
\end{equation}
(cf. \eqref{dualityElements}).
\end{proposition} 
%\begin{equation}
%	\widetilde{\Omega}=\mathbb{N}_{0}^{N}\times  \bigtimes_{x\in V} \Omega_{x}\times \mathbb{N}_{0}^{N}
%\end{equation}
%We associate an ``extra site'', denoted by $0$ and $L+1$ attached to $1$ and $L$ respectively,
%where dual particles will accumulate in the course of time. 
%and we denote its configurations by $\bm{\xi} \in \widetilde\Omega$.
%\begin{equation}
%	\bm{\xi}=\left(\xi_{1}^{0},\ldots,\xi_{N}^{0}\right)\otimes\left(\xi_{1}^{x},\ldots,\xi_{N}^{x}\right)^{\otimes L}\otimes \left(\xi_{1}^{L+1},\ldots,\xi_{N}^{L+1}\right)
%\end{equation}
%where the component $\xi_{a}^{x}$ is interpreted as the number of dual particles of type $a\in \{1,\ldots,N\}$ at site $x$, 
%and the component $\xi_{a}^{0},\xi_{a}^{L+1}$  give the number at the extra sites. \\
%where we denote the \textit{density} of the species $a\in \{2,\ldots,N\}$ imposed by the reservoir at site $x\in V$  by 
%\begin{equation}
%	\rho_{a}^{x}:=\frac{\alpha_{a}^{x}}{|\alpha^{x}|}
%\end{equation}

\paragraph{Proof of Proposition~\ref{propositin-duality-RD}:}
we only prove duality for the bulk generator, since the boundary generators are the same of the multispecies stirring process. In the spirit of Section~\ref{sectionDuality}, it is convenient to write the Hamiltonian of the process, that is defined through the basis elements \eqref{actionE} of the ${gl}(N)$ Lie algebra as 
\begin{equation}
	H^{rd}=H_{\text{left}}+\sum_{x=1}^{L-1}\left(\sigma_{11}\mathcal{H}_{x,x+1}+\sigma_{12}\sum_{c=1}^{N-2}\mathcal{H}_{x,x+1}^{c}+(\Upsilon-2\sigma_{12})\mathcal{H}_{x,x+1}^{m}\right)+H_{\text{right}}\dt
\end{equation}
Here, $\mathcal{H}_{x,x+1}$ is  \eqref{edgeHamiltonian}  and $H_{\text{left}},H_{\text{right}}$ is \eqref{siteHamiltonian} (with the choice of the boundary parameters done in \eqref{boundaryParamRD}). We, then have that 
\begin{equation}
	\mathcal{H}_{x,x+1}^{c}=\sum_{A,B=1}^{N}\left(E_{h_{c}(A)B}\otimes E_{h_{c}(B)A}-E_{AA}\otimes E_{BB}\right)
\end{equation}
and 
\begin{equation}
	\mathcal{H}_{x,x+1}^{m}=\sum_{A,B=1}^{N}\left(E_{BA}\otimes \mathbbm{1}-E_{AA}\otimes \mathbbm{1}\right)\dt
\end{equation}
We prove that, for arbitrary $x\in\{1,\ldots,L\}$,
\begin{align}
	&(\mathcal{H}_{x,x+1}^{c})^{T}D=D\mathcal{H}_{x,x+1}^{c}\qquad \forall c\in \{1,\ldots,N-2\}\label{cDualityRelation}\\
	&(\mathcal{H}_{x,x+1}^{m})^{T}D=D\mathcal{H}_{x,x+1}^{m}\label{mDualityRelation}\dt
\end{align}
Since the matrix $R_{x}$ is the same as in Section~\ref{sectionDuality}, we shall show that $\mathcal{H}_{x,x+1}^{c}$ and $\mathcal{H}_{x,x+1}^{m}$ commute with $\exp{(E^{x})}\exp{(E^{x+1})}$, where $E$ is defined in \eqref{EquationEx}. Using \eqref{eq:comgl}, the bilinearity and associativity of the Kronecker product and the bilinearity of the brackets we obtain 
\begin{align*}
	&\left[\left(\sum_{a=1}^{N-1}E_{aN}^{x}\otimes\sum_{a=1}^{N-1}E_{aN}^{x+1}\right), \sum_{D,K=1}^{N}\left(E_{h_{c}(K)D}^{x}\otimes E_{h_{c}(D)K}^{x+1} -E_{DD}^{x}\otimes E_{KK}^{x+1}\right)\right]
	\\=&
%	\sum_{d,k=1}^{N-1}\sum_{a,b=1}^{N}\left\{\left((E_{da}^{x}\delta_{h_{c}(b)N}-E_{h_{c}(b)N}^{x}\delta_{ad})-(E_{da}^{x}\delta_{aN}-E_{aN}^{x}\delta_{da})\right)\right. \\ \otimes&\left. \left((E_{kb}^{x+1}\delta_{h_{c}(a)N}-E_{h_{c}N}^{x+1}\delta_{kb})-(E_{kb}^{x+1}\delta_{bN}-E_{bN}^{x+1}\delta_{kb})\right)\right\}
%	\\=&
%	\sum_{d,k=1}^{N-1}\sum_{a,b=1}^{N}\left\{\left(E_{da}^{x}-E_{h_{c}(b)N}^{x}-E_{da}^{x}+E_{aN}^{x}\right)\otimes\left(E_{kb}^{x+1}-E_{h_{c}(a)N}^{x+1}-E_{kb}^{x+1}+E_{bN}^{x+1}\right)\right\}
%	\\=&
	\sum_{a,b=1}^{N-1}\sum_{D,K=1}^{N}\left\{\left(E_{DN}^{x}-E_{h_{c}(K)N}^{x}\right)\otimes \left(E_{DN}^{x+1}-E_{h_{c}(K)N}^{x+1}\right)\right\}=0\dt
\end{align*}
In the last equality we used the fact that map $h_{c}(\cdot)$ is surjective. 
%\paragraph{Duality for $H_{x,x+1}^{m}$} 
Concerning the commutator between $\mathcal{H}_{x,x+1}^{m}$ the proof is similar and gives
\begin{equation}
	\left[\sum_{a=1}^{N-1}E_{aN}^{x}\otimes \sum_{b=1}^{N-1}E_{bN}^{x},\;\sum_{D,K=1}^{N}\left(E_{K a}^{x}\otimes \mathbbm{1}^{x+1}-E_{DD}^{x}\otimes \mathbbm{1}^{x+1}\right)\right]=0\dt
\end{equation}
\begin{flushright}
	$\square$
\end{flushright}

>>>>

where $\mathcal{P}_{x_1}(1)$ is the probability that a single random walk started at $x_1$ is absorbed. Then, using \eqref{exactCorrelations}, we have

@@

@@

Therefore, we compute the third cumulant, i.e. the three-point connected correlation as 
\begin{align}
	&\mathbb{E}\left[\left(\mathbbm{1}_{\{Y_{x_{1}}=a_{1}\}}-\langle\rho_{a_{1}}^{x_{1}}\rangle \right)\left(\mathbbm{1}_{\{Y_{x_{2}}=a_{2}\}}-\langle\rho_{a_{2}}^{x_{2}}\rangle \right)\left(\mathbbm{1}_{\{Y_{x_{3}}=a_{3}\}}-\langle\rho_{a_{3}}^{x_{3}}\rangle \right)\right]\nonumber\\
	=&-\frac{2x_{1}(L+1-2x_{2})(L+1-x_{3})}{(L+1)^{3}(L-1)L}(\alpha_{a_{1}}-\beta_{a_{1}})(\alpha_{a_{2}}-\beta_{a_{2}})(\alpha_{a_{3}}-\beta_{a_{3}})\dt
\end{align} 
\begin{remark} The first and second cumulants, computed in \eqref{one-pts-corr} and \eqref{two-pts-corr} respectively, match with those found in Section 4.3 of \cite{vanicat2017exact}.
	\end{remark}

@@

@@

\paragraph{Proof of formula \eqref{ResulsBasis}}  
We consider the vector with elements given by the operators of the MPA $(X_{1},\ldots,X_{N})$ and we act on it with the similarity transformation $S=S_{2}S_{1}$ obtaining new operators $(\Xt_{1},\ldots,\Xt_{N})$ that will satisfy simpler commutation relations. 
% For the sake of notation we denote by
% \begin{equation}
%     \begin{pmatrix}
% 		X_{a}\\ 
% 		X_{N}
% 	\end{pmatrix}=\begin{pmatrix}
% 	    X_{1}\\
%      \vdots\\
%      X_{N}
% 	\end{pmatrix}\co
% \end{equation}
We define the transformed matrix product operators
via $\tilde X=S X$ such that
\begin{equation}\label{Xtildes2b}
 \tilde X_a=X_{a}-\beta_{a}(X_{a}+\ldots+X_{N})\,,\qquad \tilde X_N= 
				X_{1}+\ldots +X_{N}
\end{equation} 
where  $ a\in \{1,\ldots,N-1\}$.
% 
% \begin{equation}
% 	\begin{pmatrix}
% 		\Xt_{1}\\ 
% 		\vdots\\
% 		\Xt_{N-1}\\
% 		\Xt_{N}
% 	\end{pmatrix} =S\begin{pmatrix}
% 		X_{1}\\
% 		\vdots 
% 		\\
% 		X_{N-1}
% 		\\X_{N}
% 	\end{pmatrix}=\begin{pmatrix} 
% 		X_{1}-\beta_{1}(X_{1}+\ldots+X_{N})\\
% 		\vdots \\
% 		X_{N-1}-\beta_{N-1}(X_{1}+\ldots+X_{N})\\
% 				X_{1}+\ldots +X_{N}\\
% 	\end{pmatrix}\dt
% \end{equation}
We can also reverse the transformation by $S^{-1}$ and get: 
\begin{equation}\label{Xes}
 X_{a}=	\Xt_{a}+\beta_{a}\Xt_{N}\,,\qquad
	X_{N} = \beta_N\Xt_{N}-(\Xt_{1}+\ldots+\Xt_{N-1})\,.
\end{equation}

Summing over $b\in \{1,\ldots,N\}$ in commutation relations \eqref{bulk} we have
\begin{equation} 
	\left[X_{a},\Xt_{N}\right]=(\alpha_{a}-\beta_{a})\Xt_{N}
\end{equation}
with $\widetilde{X}_{N}= {X}_{1}+\ldots+ {X}_{N}$. 
Therefore, using \eqref{Xes}
we obtain the commutation relations for $\Xt_{a}$ 
\begin{equation}\label{commutationsBula}
	\left[\Xt_{a},\Xt_{N}\right]=(\alpha_{a}-\beta_{a})\Xt_{N}\co
\end{equation} 
cf.~\eqref{commutationsBula2}.
Moreover using \eqref{Xtildes2b} and \eqref{Xes}, the action of $\Xt_{a}$ on the boundary vectors are given by 
\begin{equation}\label{commLEFT}
	 \langle \langle W|  \left((\alpha_{a}-\beta_{a})\Xt_{N}-\Xt_{a}\right)=(\alpha_{a}-\beta_{a}) \langle \langle W|  
	\co
\end{equation}
\begin{equation}\label{commRIGHT}
	\Xt_{a}  |V\rangle\rangle = (\alpha_{a}-\beta_{a}) |V\rangle\rangle 
	\co
\end{equation}  
cf.~\eqref{commLEFT2}.
Using the transformed operators $(\widetilde{X}_{a})_{a\in\{1,\ldots,N\}}$, the vector $|\Psi^{''}\rangle$ of the matrix product ansatz is written as
\begin{equation}\label{toDetermineTilde}
	|\Psi^{''}\rangle = \frac{1}{Z_{L}}\sum_{\bm{\tau}\in \Omega^{'}} \langle \langle W| \widetilde{X}_{\tau_{1}}\cdots \widetilde{X}_{\tau_{L}}
	|V\rangle\rangle \,|\bm{\tau}\rangle
\end{equation} 
where
the normalization is given by 
\begin{align}
		Z_{L}&= \langle \langle W|  (X_{1}+\ldots+X_{N})^{L} |V\rangle\rangle = \langle \langle W|  \widetilde{X}_{N}^{L} |V\rangle\rangle \dt
\end{align}

@@

@@

@@

@@

	\xi_{A}^{x}=	\begin{cases}
	\sum_{y=1}^{L}\sum_{b=1}^{N-1}\mathbbm{1}_{\{y=x_{k}\}}\mathbbm{1}_{\{b=a_{k}\}}\qquad &\text{if}\quad x=x_{k}\;\text{and} \; A=a_{k}\\
		 0\qquad &\text{otherwise}

@@

	\\&\times

@@

		\mathcal{L}_{\text{left}}^{TH}f(\bm{n})=&\sum_{r_{1}=0}^{\nu}\ldots\sum_{r_{N}=0}^{\nu}\mathbbm{1}_{\{r_{1}+\ldots+r_{N}=\nu\}}\left\{f(r,n^{2},\ldots,n^{L})-f(\bm{n})\right\}\\&\times

@@

	\todo{is the last term in the product?}

@@

\todo{Why not merge sum or is there a typo?}

@@

\subsubsection{Duality for the reaction-diffusion process}
In this section we formulate duality for the multispecies reaction diffusion process. This answer to the second question addressed at the beginning of Section~\ref{Subsection-RD}.  The dual process is defined on the enlarged state space $\widetilde{\Omega}_{L}$.  
\begin{proposition}[Duality for the reaction-diffusion process]\label{propositin-duality-RD}
The reaction-diffusion multispecies stirring process $(\bm{n}(t))_{t\geq 0}$, on the state space $\Omega_{L}$, with generator $\mathcal{L}^{rd}$ defined in \eqref{RDGenerator} is dual to the process $(\bm{\xi}(t))_{t\geq 0}$ on the state space $\widetilde{\Omega}_{L}$ with dual generator
\begin{equation}\label{DualGeneratorRD}
	\widetilde{\mathcal{L}}_{\text{left}}^{rd}=\widetilde{\mathcal{L}}^{rd}+\sum_{x=1}^{L-1}\mathcal{L}_{x,x+1}^{rd}+\widetilde{\mathcal{L}}_{\text{right}}^{rd}
\end{equation}
where 
$\mathcal{L}_{x,x+1}^{rd}$ is defined in \eqref{edgeGeneratorRD} and, $\widetilde{\mathcal{L}}_{\text{left}}^{rd}$ and $\widetilde{\mathcal{L}}_{\text{right}}^{rd}$ are the absorbing dual generators $\widetilde{\mathcal{L}}_{x}$ defined in \eqref{siteDualGenerator} and acting at sites $1$ and $L$, with the specific choice of the reservoirs parameters \eqref{boundaryParamRD}. The duality function
is the same of the stirring process, i.e.
\begin{equation}
	D(\bm{n},\bm{\xi})=\left(\prod_{a=1}^{N-1}\left(\rho_{a}^{\text{left}}\right)^{\xi_{a}^{0}}\right)\left(\prod_{x=1}^{L}\frac{(\nu -\sum_{a=1}^{N-1}\xi_{a}^{x})!}{\nu!}\prod_{a=1}^{N-1}\frac{n_{a}^{x}!}{(n_{a}^{x}-\xi_{a}^{x})!}\right)\left(\prod_{a=1}^{N-1}\left(\rho_{a}^{\text{right}}\right)^{\xi_{a}^{L+1}}\right)\co
\end{equation}
cf. \eqref{dualityElements}.
\end{proposition} 
%\begin{equation}
%	\widetilde{\Omega}=\mathbb{N}_{0}^{N}\times  \bigtimes_{x\in V} \Omega_{x}\times \mathbb{N}_{0}^{N}
%\end{equation}
%We associate an ``extra site'', denoted by $0$ and $L+1$ attached to $1$ and $L$ respectively,
%where dual particles will accumulate in the course of time. 
%and we denote its configurations by $\bm{\xi} \in \widetilde\Omega$.
%\begin{equation}
%	\bm{\xi}=\left(\xi_{1}^{0},\ldots,\xi_{N}^{0}\right)\otimes\left(\xi_{1}^{x},\ldots,\xi_{N}^{x}\right)^{\otimes L}\otimes \left(\xi_{1}^{L+1},\ldots,\xi_{N}^{L+1}\right)
%\end{equation}
%where the component $\xi_{a}^{x}$ is interpreted as the number of dual particles of type $a\in \{1,\ldots,N\}$ at site $x$, 
%and the component $\xi_{a}^{0},\xi_{a}^{L+1}$  give the number at the extra sites. \\
%where we denote the \textit{density} of the species $a\in \{2,\ldots,N\}$ imposed by the reservoir at site $x\in V$  by 
%\begin{equation}
%	\rho_{a}^{x}:=\frac{\alpha_{a}^{x}}{|\alpha^{x}|}
%\end{equation}

\paragraph{Proof of Proposition~\ref{propositin-duality-RD}:}
we only prove duality for the bulk generator, since the boundary generators are the same of the multispecies stirring process. In the spirit of Section~\ref{sectionDuality}, it is convenient to write the Hamiltonian of the process, that is defined through the basis elements \eqref{actionE} of the ${gl}(N)$ Lie algebra as 
\begin{equation}
	H^{rd}=H_{\text{left}}+\sum_{x=1}^{L-1}\left(\sigma_{11}\mathcal{H}_{x,x+1}+\sigma_{12}\sum_{c=1}^{N-2}\mathcal{H}_{x,x+1}^{c}+(\Upsilon-2\sigma_{12})\mathcal{H}_{x,x+1}^{m}\right)+H_{\text{right}}\dt
\end{equation}
Here, $\mathcal{H}_{x,x+1}$ is  \eqref{edgeHamiltonian}  and $H_{\text{left}},H_{\text{right}}$ is \eqref{siteHamiltonian} (with the choice of the boundary parameters done in \eqref{boundaryParamRD}). We, then have that 
\begin{equation}
	\mathcal{H}_{x,x+1}^{c}=\sum_{A,B=1}^{N}\left(E_{h_{c}(A)B}\otimes E_{h_{c}(B)A}-E_{AA}\otimes E_{BB}\right)
\end{equation}
and 
\begin{equation}
	\mathcal{H}_{x,x+1}^{m}=\sum_{A,B=1}^{N}\left(E_{BA}\otimes \mathbbm{1}-E_{AA}\otimes \mathbbm{1}\right)\dt
\end{equation}
We prove that, for arbitrary $x\in\{1,\ldots,L\}$,
\begin{align}
	&(\mathcal{H}_{x,x+1}^{c})^{T}D=D\mathcal{H}_{x,x+1}^{c}\qquad \forall c\in \{1,\ldots,N-2\}\label{cDualityRelation}\\
	&(\mathcal{H}_{x,x+1}^{m})^{T}D=D\mathcal{H}_{x,x+1}^{m}\label{mDualityRelation}\dt
\end{align}
Since the matrix $R_{x}$ is the same as in Section~\ref{sectionDuality}, we shall show that $\mathcal{H}_{x,x+1}^{c}$ and $\mathcal{H}_{x,x+1}^{m}$ commute with $\exp{(E^{x})}\exp{(E^{x+1})}$, where $E$ is defined in \eqref{EquationEx}. Using \eqref{eq:comgl}, the bilinearity and associativity of the Kronecker product and the bilinearity of the brackets we obtain 
\begin{align*}
	&\left[\left(\sum_{a=1}^{N-1}E_{aN}^{x}\otimes\sum_{a=1}^{N-1}E_{aN}^{x+1}\right), \sum_{D,K=1}^{N}\left(E_{h_{c}(K)D}^{x}\otimes E_{h_{c}(D)K}^{x+1} -E_{DD}^{x}\otimes E_{KK}^{x+1}\right)\right]
	\\=&
%	\sum_{d,k=1}^{N-1}\sum_{a,b=1}^{N}\left\{\left((E_{da}^{x}\delta_{h_{c}(b)N}-E_{h_{c}(b)N}^{x}\delta_{ad})-(E_{da}^{x}\delta_{aN}-E_{aN}^{x}\delta_{da})\right)\right. \\ \otimes&\left. \left((E_{kb}^{x+1}\delta_{h_{c}(a)N}-E_{h_{c}N}^{x+1}\delta_{kb})-(E_{kb}^{x+1}\delta_{bN}-E_{bN}^{x+1}\delta_{kb})\right)\right\}
%	\\=&
%	\sum_{d,k=1}^{N-1}\sum_{a,b=1}^{N}\left\{\left(E_{da}^{x}-E_{h_{c}(b)N}^{x}-E_{da}^{x}+E_{aN}^{x}\right)\otimes\left(E_{kb}^{x+1}-E_{h_{c}(a)N}^{x+1}-E_{kb}^{x+1}+E_{bN}^{x+1}\right)\right\}
%	\\=&
	\sum_{a,b=1}^{N-1}\sum_{D,K=1}^{N}\left\{\left(E_{DN}^{x}-E_{h_{c}(K)N}^{x}\right)\otimes \left(E_{DN}^{x+1}-E_{h_{c}(K)N}^{x+1}\right)\right\}=0\dt
\end{align*}
In the last equality we used the fact that map $h_{c}(\cdot)$ is surjective. 
%\paragraph{Duality for $H_{x,x+1}^{m}$} 
Concerning the commutator between $\mathcal{H}_{x,x+1}^{m}$ the proof is similar and gives
\begin{equation}
	\left[\sum_{a=1}^{N-1}E_{aN}^{x}\otimes \sum_{b=1}^{N-1}E_{bN}^{x},\;\sum_{D,K=1}^{N}\left(E_{K a}^{x}\otimes \mathbbm{1}^{x+1}-E_{DD}^{x}\otimes \mathbbm{1}^{x+1}\right)\right]=0\dt
\end{equation}
\todo{shall we remove all site indices given that there is the tensor product?}
\begin{flushright}
	$\square$
\end{flushright}

\appendix
\section*{Appendix}
\section{Interpretation of $\Psi^{'}$ as marginals of the steady state distribution}\label{subsection-interpretationPSIp}
\todo{Why do we never refer to this appendix?}
We still miss the interpretation of $\Psi^{'}(\bm{\tau})$ when  $\bm{\tau}= (N,\ldots,N)$. We show here that for this particular configuration
$\Psi^{'}(N,\ldots,N) =1$. Indeed, for any $\bm{\tau}\in \Omega^{'}$ we have that 
\begin{align}
	\Psi^{'}(\bm{\tau})=&\langle\bm{\tau}|\Psi^{'}\rangle=\sum_{\bm{\sigma}\in \Omega^{'}}\langle \bm{\tau}|S_{1}|\bm{\sigma}\rangle\langle \bm{\sigma}|\Psi\rangle=\sum_{\bm{\sigma}\in \Omega^{'}}\langle \bm{\sigma}|S_{1}^{T}|\bm{\tau}\rangle\langle \bm{\sigma}|\Psi\rangle\dt
\end{align} 
We introduce the dual configuration $\bm{\xi}(\bm{\tau})=(N,\tau_{1},\ldots,\tau_{L},N)$ where two extra-sites occupied by the species $N$ have been attached to the vector $\bm{\tau}$, transforming it from a vector of dimension $L$ to one of dimension $L+2$. More explicitly, this dual configuration is given by
\begin{equation}
	\begin{cases}
		\xi_{a}^{x}(\bm{\tau})= \delta_{\tau_{x},a}\qquad \forall a\in \{1,\ldots N\}\quad \forall x\in \{1,\ldots,L\}\\
		\xi_{N}^{0}(\bm{\tau})=\xi_{N}^{L+1}(\bm{\tau})=1
	\end{cases}\dt
\end{equation} Now we consider the elements $\langle\bm{\sigma}|D|\xi(\bm{\tau})\rangle$ of the duality matrix $D$ and we have
\begin{align}\label{useful-InInterpretation}
	\Psi^{'}(\bm{\tau})=&\sum_{\bm{\sigma}\in \Omega^{'}}\langle \bm{\sigma}|D|\bm{\xi}(\bm{\tau})\rangle\langle \bm{\sigma}|\Psi\rangle\dt
\end{align}
We observe that for any $x\in \{1,\ldots,L\}$
\begin{equation}
	\prod_{a=1}^{N-1}\mathbbm{1}_{\{\delta_{\sigma_{x},a}\geq\xi_{a}^{x}(\bm{\tau})\}}=\left(\mathbbm{1}_{\{\sigma_{x}=\tau_{x}\}}\right)^{(1-\delta_{\tau_{x},N})}%=\begin{cases}
		%	\mathbbm{1}_{\{\sigma_{x}=\tau_{x}\}}\qquad &\text{if}\quad \tau_{x}\neq N\\
		%	1\qquad  &\text{if}\quad \tau_{x}= N
		%	\end{cases}\dt
\end{equation}
Indeed, we have that 
\begin{equation}
	\prod_{a=1}^{N-1}\mathbbm{1}_{\{\delta_{\sigma_{x},a}\geq\xi_{a}^{x}(\bm{\tau})\}}=\begin{cases}
		\mathbbm{1}_{\{\sigma_{x}=\tau_{x}\}}\qquad &\text{if}\quad \tau_{x}\neq N\\
		1\qquad  &\text{if}\quad \tau_{x}= N
	\end{cases}\dt
\end{equation}
Therefore using the duality function \eqref{ElementsDualityMatrixChain_1} and equation~\ref{useful-InInterpretation}, we obtain 
\begin{equation}\label{interpretation-PsiP}
	\Psi^{'}(\bm{\tau})=\mathbb{E}\left[\prod_{x=1}^{L}\left(\mathbbm{1}_{\{Y_{x}=\tau_{x}\}} \right)^{(1-\delta_{\tau_{x},N})}\right]\dt
\end{equation} 
This last equation completes and generalizes \eqref{corr-psi1} by showing that $\Psi^{'}(N,\ldots,N) =1$.
